---
title: "p8105_hw3_mnw2132"
author: "Mary Williams"
date: "10-16-2024"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r load_libraries}
library(tidyverse)
library(ggplot2)
```

### Problem 1

```{r}
library(p8105.datasets)
data("ny_noaa")
```

When looking at the data with: 
```{r}
str(ny_noaa)
summary(ny_noaa)
sum(is.na(ny_noaa))
```
there are 2,595,176 rows with 7 variables including id, date, prcp, snow, snwd, tmax and tmin. id, tmax, and tmin are charecters, prcp, snow and snwd are integers, and date is just date. Thre are 3387623 missing variables. The rows that have all information besides id and date missing, would be good to remove. 

# Data cleaning
*Separate year, month, and date; clean names; make tmin and tmax numeric, and change prcp, tmax and tmin, to no longer be "a tenth of..."*
```{r  clean_data}
cleaned_noaa = ny_noaa %>%
  janitor::clean_names() %>%
  separate("date", into = c("year", "month", "day"), sep = "3") %>%
   mutate(
    tmin = as.numeric(tmin),
    tmax = as.numeric(tmax),
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10
  )
```
# Checking cleaned data 
```{r}
summary(cleaned_noaa)
str(cleaned_noaa)
cleaned_noaa
```

# Most common observed snowfall values
```{r common_observed_snowfall}
cleaned_noaa %>%
group_by(snow) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```
The most common snowfall values were "0" with 2008508 and "NA" with 381221. This is likely because it does not snow often, especially in some of the locations observed. 

# Max temperature for January and July
```{r}
cleaned_noaa %>%
  mutate(
    month = as.integer(month)) %>%
    filter(month %in% c(-01, -07)) %>%
  ggplot(aes(x = year, y = tmax, color = id)) +
  geom_point() +
  labs(title = "Max Temp in January and July",
       x = "Year",
       y = "Max temp (C)")


unique(cleaned_noaa$month)
unique(cleaned_noaa$day)

cleaned_noaa %>%
  mutate(
    date = as.Date(date),  # Ensure date is in Date format
    month = as.integer(format(date, "%m"))  # Extract month as integer
  ) %>%
  filter(month %in% c(1, 7)) %>%  # Filter for January and July
  ggplot(aes(x = year, y = tmax, color = id)) +
  geom_point() +
  labs(title = "Max Temp in January and July",
       x = "Year",
       y = "Max temp (°C)")

```
